ICT and Future Productivity:\\Evidence and Theory of a GPT\thanks{Correspondence: Department of Economics, Boston College, 140 Commonwealth Avenue, Chestnut Hill, MA 02467. Email: brianti@bc.edu (Marco Brianti) and gati@bc.edu (Laura G\'ati).}

### 2018-07-26 Latex/Draft_Summer2018/Summer2018.tex
WRITE WRITE WRITE

Although our results are robust over different specifications, an important critique to our empirical strategy is reverse causality coming from news on future TFP. As suggested by the news-shock literature, the positive reaction of ICTI on impact may be triggered by signals related future increases in TFP and not to contemporaneous ICT technological improvements. In other words, our identification strategy may confound our shock of interest with a news shock which contemporaneously enhances investment in ICT capital goods. We address this issue by providing a series of alternative identification strategies which we show deliver the same time series of ICT innovations as our initial specification. The heart of these robustness checks is sequential identification of news and ICT shocks: we firstly identify a news shock in the spirit of \cite{barsky2011news}, and subsequently we identify our sectoral ICT shock using the previous identification strategy. Encouragingly, controlling for signals regarding future movements in TFP does not affect our results. We view this as strengthening our statement relating movements in current ICT technology to future TFP.

@@

The model presented has several common features with the two-sector environment of \cite{greenwood1997long} where one sector displays a faster technology than the other. Three main assumptions distinguish our model from the one by \cite{greenwood1997long}. Firs of all, we focus specifically on the ICT capital sector, rather than equipment, as the sector with faster technological improvement. This assumption will be necessary if we want to fit the fact that relative prices between ICT investment and the overall economy display a downward sloping trend.\footnote{In addition, we also rely on the results of \cite{whelan2003two} and \cite{oulton2007investment} to correctly define productivity and output share in our setting.} A second difference is that we consider two production functions - one for each sector. As \cite{oulton2007investment} shows, the two formulations are isomorphic, so our modeling choice has no implication for the resulting dynamics. The reason we choose this approach is because it allows for definitions of GDP and TFP in the model that are consistent with the way these variables are constructed in the national income accounts. Finally, we assume that the total ICT capital stock has a positive spillover effect on the production function of both sectors. The general-purpose nature of ICT capital works through its own diffusion because information sharing improves only when new communication technologies are widely spread over the population.

@@

%%%%%%%%%%%%%%%%%%%%               MODEL                 %%%%%%%%%%%%%%%%%% 
\section{Experiments}\label{section:experiments}

@@

>>>>

Although our results are robust over different specifications, an important critique to our empirical strategy is reverse causality coming from news on future TFP. As suggested by the news-shock literature, the positive reaction of ICTI on impact may be triggered by signals related future increases in TFP and not to contemporaneous ICT technological improvements. In other words, our identification strategy may confound our shock of interest with a news shock which contemporaneously enhances investment in ICT capital goods. We address this issue by providing a series of alternative identification strategies which we show deliver the same time series of ICT innovations as our initial specification. The heart of these robustness checks is sequential identification of news and ICT shocks: we firstly identify a news shock in the spirit of \cite{barsky2011news}, and subsequently we identify our sectoral ICT shock using the previous identification strategy. Encouragingly, controlling for signals regarding future movements in TFP does not affect our results. We view this as strengthening the causality relation from current ICT investment to future TFP.

@@

The model presented has several common features with the two-sector environment of \cite{greenwood1997long} where one sector displays a faster technology than the other. Three main assumptions distinguish our model from the one by \cite{greenwood1997long}. First of all, we focus specifically on the ICT capital sector, rather than equipment, as the sector with faster technological improvement. This assumption will be necessary if we want to fit the fact that relative prices between ICT investment and the overall economy display a downward sloping trend.\footnote{In addition, we also rely on the results of \cite{whelan2003two} and \cite{oulton2007investment} to correctly define productivity and output share in our setting.} A second difference is that we consider two production functions - one for each sector. As \cite{oulton2007investment} shows, the two formulations are isomorphic, so our modeling choice has no implication for the resulting dynamics. The reason we choose this approach is because it allows for definitions of GDP and TFP in the model that are consistent with the way these variables are constructed in the national income accounts. Finally, we assume that the total ICT capital stock has a positive spillover effect on the production function of both sectors. The general-purpose nature of ICT capital works through its own diffusion because information sharing improves only when new communication technologies are widely spread over the population.

@@

Although both the model and impulse-response matching exercise should be considered as preliminary, we strongly believe they are already able to rationalize empirical results obtained in Section \ref{section:empirics} and convey the main message of this project. Formally, the question we tried to answer is why a contemporaneous technological shock specific to the ICT sector is able to trigger increases in future total factor productivity. In other words, why an ICT-investment shock is able to increase productivity of ICT goods users with some lags.

We interpret ICT capital as being the general-purpose technology of the U.S. economy over the last $3$ decades. Specifically, when an ICT technology shock hits the economy, both sectors invest more on ICT capital goods since they are now cheaper to purchase and easier to produce. Now, in order to link current higher investment on ICT durable goods to future TFP, we assume that both production functions are augmented by a spillover effect related to the overall stock of ICT capital which is not internalized by the agents.\footnote{This externality is governed by the parameter $\gamma$ which has been estimated via impulse-response matching to be robustly larger than zero. Using different empirical results, changing the initial calibration, or the methodology do not qualitative change the estimated value of $\gamma$ close to $0.6$.} These assumptions are mechanically suitable for two main reasons. First of all, this spillover enters into the estimated Solow residual since it takes the form of an externality which implies that industries are not paying for this additional benefit. Moreover, since this spillover arises from a state variable, it leads to a dynamics where the accumulation of ICT durable goods is a slow process and the benefits of an ICT technology shock show up in the productivity of ICT-capital users with some lags and this effect tends to be persistent.

Although this assumption is mathematically suitable to track empirical impulse responses, we consider important to highlight that this positive externality is sustained by an intuitive and reliable economic intuition which distinguishes the general-purpose nature of ICT capital. Since the purpose of ICT goods is to enhance communication, the quality and speed of information sharing is also determined by the overall diffusion of these goods. As a simple example, assume that in 1995 a scholar decides to purchase a computer with an Internet subscription in order to speed up the acquisition of information. She is very satisfied of this investment and does not regret the paid amount. However, at the same time, a former colleague who is now working in a different university decides to buy a computer with an Internet subscription as well. Once she learns of this development, she also decides to exploit her computer to exchange research material with her former colleague. They eventually decide to start a new project together since being in two different universities is not anymore a constraint for the quality of their joint research. Both scholars did not pay for this additional benefit since they did not expect to have simultaneously a computer with an Internet subscription. However, what is actually happening is that their research production has been boosted by the positive externality that also the other colleague is endowed with the same technology.

As a result, we expect that the nature of ICT capital should display two main features. First of all, the effectiveness of an ICT good must be related to its own diffusion as pointed out by the spillover effect. Moreover, the cost of ICT items should not initially reflect an unexpected increase in their diffusion as pointed out by the assumption that the first feature is not internalized.

%%%%%%%%%%%%%%%%%%%%            EXPERIMENTS             %%%%%%%%%%%%%%%%%% 
%\section{Experiments}\label{section:experiments}

%TBD. %I personally would wait after some comments/conferences

@@

In the first part of the paper, we investigate the effect of technological shocks specific to the ICT sector on future TFP. Using Structural VAR techniques on quarterly aggregate U.S. data from 1989, we impose two identifying restrictions to correctly identify ICT shocks: (i) they must be orthogonal to current productivity and (ii) maximize the impact response of ICT investment. In response to an ICT shock, ICT investment increases on impact and stays significant for more than 10 years; relative prices persistently decline as an additional support to the reliability of our identification strategy; and TFP rises after three quarters and remains significant for more than 10-years. In addition, variance decomposition analysis suggests that the quantitative impact of ICT shocks on TFP is economically remarkable since they explain about one third of the overall TFP fluctuations over a 10-year horizon.

In order to check whether our identification assumptions may confound our shock with a news shock which contemporaneously boosts ICT investment, we provide a series of different identification strategies which deliver analogous results. Our main robustness check is a sequential identification of news shocks first and then ICT shocks which encouragingly does not affect estimates. We view this as confirmation of the causality relation which goes from current ICT investment to future TFP.

In the second part of the paper, we rationalize our empirical results by analyze a two-sector DSGE model \`a la \cite{greenwood1997long} augmented by three main assumptions. Firstly, the sector with the faster technological improvement produces specifically ICT capital. Moreover, we consider two production functions in order to have a consistent definition of GDP and TFP between the model and the data. Finally, we augment both production functions with a spillover effect related to the overall level of ICT capital which is not internalized by the agents. In particular, this last assumption points out our interpretation of the general-purpose nature of ICT capital: the effectiveness of ICT capital is enhanced by its diffusion and its cost does not reflect this indirect benefit. We use both empirical and theoretical results to estimate main parameters of the model via impulse-response matching to an ICT technology shock. Results point out that the parameter which governs the size of the spillover effect is positive, suggesting that our interpretation of the general-purpose nature of ICT capital is confirmed by the data.
